\section{OSOR}
\label{sec:osor}

\par \href{https://joinup.ec.europa.eu/page/osor.eu}{Open Source Observatory Repository} - Free Libre Open Source Software \textit{- FLOSS -} Forge oriented to European Public Administration. Giving a technic base and guidelines to develop FLOSS projects throught good practices. Projects has to be licensed with \href{http://joinup.ec.europa.eu/software/page/eupl/licence-eupl}{EUPL} European Union Public License. \textit{Only} Public Administrations funded FLOSS project can be hosted in OSOR you can not advertise a project by yourself.

\par The project started in 2007 and in December 2011 migrated to \href{https://joinup.ec.europa.eu/page/osor.eu}{JoinUP} the new development platform.

\subsection{Development Tools}

\par OSOR main goal is to create Communities for each FLOSS project thus create an environment in a common place to all Public Administration developments.

\begin{itemize}
	\item \textit{\href{http://gforge.com/gf/}{Gforge}} - Base for the FLOSS Forge that became Fusion Forge in 2009 (http://blog.bitergia.com/2012/11/16/the-history-of-fusionforge-and-gforge/). Strong communication with Gforge community.
	\item \textit{Repository} - Apache \href{http://subversion.apache.org/}{Subversion} for source code management.
	\item \textit{CMS -} Service for users \href{http://plone.org/}{Plone CMS}
	\item \textit{Mailing list} - \href{http://www.gnu.org/software/mailman/}{GNU Mailman}.
\end{itemize}\textit{I couldn't found more information about development tools for the project.}

\subsection{How to contribute}

\par The project ended in December 2011 so is not possible to collaborate in this project, there is no community supporting the project. Is a FLOSS project thus everyone could continue its development or play with the code to learn.

\par This summarizes community management ROLES big picture inside OSOR project A.K.A. \textit{Cheerleaders}:

\begin{itemize}
	\item \textbf{Cheerleadears} - People dedicated to spread OSOR and advertise the goodnesses of using OSOR.
\begin{itemize}
	\item \textit{News} - \href{https://twitter.com/Sjig}{Gijs Hillenius} - Journalist specialized in FLOSS and communication with other FLOSS communities, promoting workshops, events and meetings between communities.
	\item \textit{Coordination} - \href{http://www.unisys.com/unisys/}{Unisys} Ismael Olea (\href{https://twitter.com/olea}{@olea}) spreading OSOR use in Europe.
	\item \textit{Dynamizer} - \href{http://www.gopa-cartermill.com/}{GOPA}
\end{itemize}
\end{itemize}

% section osor (end)